---
title: "ParlGov data"
author: "Holger Doering --- holger.doering@gesis.org"
date: "`r format(Sys.time(), '%d %B %Y, %H:%M')`"
output:
  html_notebook:
    # code_folding: hide
    toc: yes
    toc_float: yes
---
<style type="text/css"> <!-- .table { width: auto } ---> </style>


Document created: `r format(Sys.Date(), "%d %b %Y")`

__Examples to access ParlGov data with Tidyverse-R__


## One table

Download (if not available) and read main ParlGov table (parties, elections, cabinets)

```{r single-table, message=FALSE}
library(tidyverse)

tbl <- "view_cabinet.csv"  # view_party.csv -- view_election.csv -- view_cabinet.csv
if( ! tbl %in% list.files()) {
  download.file(paste0("https://parlgov.org/data/parlgov-development_csv-utf-8/", tbl),  tbl)
}
dt_raw <- read_csv(tbl)
```


## Multiple tables

Read and assign ParlGov tables — _Note:_ chunk not run (`eval=False`) to suppress output

```{r tables, eval=FALSE}
library(tidyverse)

get_parlgov_data <- function(tbl_name, obj_name = "dt") {
  url <- "https://parlgov.org/data/parlgov-development_csv-utf-8/"
  if( ! tbl_name %in% list.files()) {
    download.file(paste0(url, tbl_name),  tbl_name)
  }
  assign(obj_name, read_csv(tbl_name), envir = .GlobalEnv)
}

# get_parlgov_data("view_party.csv", "party")

pg_tbls <- c(
  party = "view_party.csv",
  elec = "view_election.csv",
  cab = "view_cabinet.csv"
  )

map2(pg_tbls, names(pg_tbls), get_parlgov_data)
```


## Database

Access data in ParlGov database

```{r database, message=FALSE}
library(tidyverse)

## Get and read cabinet and party data from ParlGov database

db_file <- "parlgov-development.db"
url <- "https://parlgov.org/data/"

# download ParlGov database if not in local folder
if( ! db_file %in% list.files()) {
  download.file(paste0(url, db_file), db_file, mode = "wb")
}

# retrieve data from ParlGov database tables
# http://www.parlgov.org/data/table/
parlgov_db <- DBI::dbConnect(RSQLite::SQLite(), db_file)
get_parlgov_table <- function(table_name) tbl(parlgov_db, table_name) |> collect()

party_raw <- get_parlgov_table("view_party")
elec_raw <- get_parlgov_table("view_election")
cab_raw <- get_parlgov_table("view_cabinet")

DBI::dbDisconnect(parlgov_db)
```


## Explore

See chunk "database" for reading data

```{r explore, message=FALSE}
library(skimr)

skim(party_raw, left_right:eu_anti_pro) |> 
  knitr::knit_print()
```
