---
title: "Small party premiers"
author: "Holger Doering --- doering@uni-bremen.de"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_float: yes
---
<style type="text/css"> <!-- .table { width: auto } ---> </style>

```{r options, include=FALSE}
knitr::opts_chunk$set(
  # results = "hide",
  message = FALSE,
  warning = FALSE,
  package.startup.message = FALSE
  )

options(
  readr.num_columns = 0,
  knitr.kable.NA = "",
  width = 100,
  tidyverse.quiet = TRUE
)
```

```{r corepacks}
library(tidyverse)
library(knitr)

ggplot2::theme_set(theme_minimal())
```

```{r}
tbl <- "view_cabinet.csv"  # view_party.csv -- view_election.csv -- view_cabinet.csv
if( ! tbl %in% list.files()) {
  download.file(paste0("http://www.parlgov.org/static/data/development-utf-8/", tbl),  tbl)
}
cab_raw <- read_csv(tbl)
```

### Question

[Fruits and votes on 5 Feb. 2020](https://fruitsandvotes.wordpress.com/2020/02/05/thuringia-leader-of-smallest-party-in-state-parliament-elected-premier-with-afd-support/)

> I don’t know how common a government led by the sixth largest party in parliament is, but I am guessing pretty uncommon. 


### ParlGov data

[ParlGov cabinet table](http://www.parlgov.org/data/table/view_cabinet/)

Prime minister parties and their size rank in parliament

Size rank determined by seat share in parliament at instance of cabinet formation.

Excluded parties and cabinets:

+ caretaker cabinets
+ parties with no seats
+ PMs with no party affiliation

```{r}
cab_pm <- 
  cab_raw %>% 
  filter(caretaker == 0, seats > 0, party_name_short != "none") %>% 
  arrange(country_name_short, start_date, desc(seats)) %>% 
  group_by(cabinet_id) %>% 
  mutate(
    size_rank = row_number(),
    size = round(100 * seats / election_seats_total, 1),
    smallest = if_else(size_rank == max(size_rank), 1, 0),
    year = lubridate::year(start_date)
    ) %>% 
  ungroup() %>% 
  filter(prime_minister == 1) %>% 
  select(
    country = country_name_short,
    year,
    date = start_date,
    cabinet = cabinet_name,
    party = party_name_short,
    seats, size, size_rank, smallest
    )
```


__`r nrow(cab_pm)`__ cabinets in __`r length(unique(cab_pm$country))`__ countries


### PM parties by size rank {.tabset .tabset-pills}

#### Table

```{r}
cab_pm %>% 
  count(size_rank) %>% 
  mutate(share = round(100 * n / sum(n), 1)) %>% 
  knitr::kable()
```

#### Figure

```{r}
pl_dt <- 
  cab_pm %>% 
  mutate(
    size_rank_5 = ifelse(size_rank < 5, size_rank, 5),
    size_rank = as_factor(size_rank_5) %>% fct_rev(),
    size_rank = fct_recode(size_rank, `> 4` = "5")
    ) 
  
ggplot(pl_dt, aes(x = size_rank)) +
  geom_bar() +
  coord_flip() +
  labs(x = NULL, y = "Frequency", fill = "") +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_bw()
```

### PM smallest party

```{r}
cab_pm %>% 
  filter(smallest == 1) %>% 
  select(-date, -smallest) %>% 
  knitr::kable()
```


### Country information

```{r}
pl <- 
  ggplot(pl_dt, aes(x = date, y = size_rank_5)) +
  geom_point(colour = "deepskyblue3", alpha = 0.5) +
  facet_wrap(~ country)

print(pl)
ggsave("pm-party-size-rank.png", pl, width = 8, height = 6)
```

```{r}
cab_pm %>% 
  arrange(desc(size_rank), size) %>% 
  select(-year, -smallest) %>% 
  DT::datatable()
```

